\textsf{Are large language models superhuman chemists?}

### 2024-03-30 src/tex/appendix.tex
style updates

\subsection{Influence of model scale}
To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
Interestingly, we find that the 7B and 70B models perform comparably, with the 7B showing lower performance ().

@@

We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

@@

Additionally, we prompted users to provide additional information about their experience in chemistry. 
While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

>>>>

\subsection{Influence of model scale}
To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
Interestingly, we find that the 7B and 70B models perform comparably, with the 7B showing lower performance (fraction of correct answers for the 7B, 13B, and 70B models are \variable{output/llama/llama_7b.txt}, \variable{output/llama/llama_13b.txt}, \variable{output/llama/llama_70b.txt}).

@@

We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

@@

Additionally, we prompted users to provide additional information about their experience in chemistry. 
While we recorded fine-grained information, e.g., their specialization, we focused on the number of years since the first university-level chemistry course.
\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

### 2024-03-30 src/tex/ms.tex
style updates

    We curated more than 7,000 question-answer pairs for a wide array of chemical sciences subfields, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average. 
    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.

    These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to improve safe and useful LLMs systematically.
\end{abstract}

@@

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones. 
Yet, it is important to recognize that while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools has been available for years and requires access to controlled facilities such as laboratories.
Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who critically reflect on every output such models produce. 
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
However, even for experts, chemical understanding and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to expert (human) chemists.

@@

To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.  %and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.

@@

We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
If the model would answer questions based on reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}

@@

\begin{figure}[!h]
    \centering
    \includegraphics{figures/confidence_vs_performance_overall.pdf}
    \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model. We prompted the models to return a confidence score on an ordinal scale to obtain those estimates. We then plot the correctness of the answers (which is calculated as the mean value of all answers being either completely correct (1) or completely incorrect (0)) against the confidence score.
        The stripplots show the individual data points, a strong density indicating a higher number of points. The lines show the mean, and the error bars show the standard error of the mean. The figure shows that the reliability of the confidence estimates varies widely across models. While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend.}
    \label{fig:confidence_vs_performance}
    \script{joint_analysis_confidence_performance.py}
\end{figure}

In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading for the questions about \gls{ghs} pictograms with an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

@@

\paragraph{Parsing}
Our parsing workflow is multistep and primarily based on regular expressions.
In the case of instruction-tuned models, we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
In the case of numbers, our regular expression was engineered to deal with various forms of scientific notation.
As initial tests indicated that models sometimes return integers in the form of words, e.g., \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., Claude 2, to parse the completion.

@@

\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b\autocite{jiang2024mixtral} Llama2 (70b),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.

@@

\paragraph{Question selection} \label{sec:subset-selection}
Since we anticipated that we would not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided to show a relevant subset of all questions to the human scorers.
For selecting the subset, we decided to address two questions:
\begin{itemize}
    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
    \item Are there areas in which the performance of humans is very different from the ones of the models?
\end{itemize}
To answer the first question, we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART. We added five random questions about the number of \gls{nmr} signals and the point group of molecules.
% https://github.com/lamalab-org/chem-bench-app/issues/92 has the details

>>>>

    We curated more than 7,000 question-answer pairs for a wide array of subfields of the chemical sciences, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average. 
    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.

    These findings underscore the dual reality that, although LLMs demonstrate remarkable proficiency in chemical tasks, further research is critical to enhancing their safety and utility in chemical sciences.
    Our findings also indicate a need for adaptations to chemistry curricula and highlight the importance of continuing to develop evaluation frameworks to improve safe and useful LLMs.
\end{abstract}

@@

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who critically reflect on every output such models produce. 
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
However, even for experts, chemical understanding and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to expert (human) chemists.

@@

To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.  %and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.

@@

We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}

@@

\begin{figure}[!h]
    \centering
    \includegraphics{figures/confidence_vs_performance_overall.pdf}
    \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model. We prompted the models to return a confidence score on an ordinal scale to obtain those estimates. We then plot the correctness of the answers (which is calculated as the mean value of all answers being either completely correct (1) or completely incorrect (0)) against the confidence score.
        The stripplots show the individual data points, a strong density indicating a higher number of points. The lines show the mean, and the error bars show the standard error of the mean. The figure shows that the reliability of the confidence estimates varies widely across models. While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend better.}
    \label{fig:confidence_vs_performance}
    \script{joint_analysis_confidence_performance.py}
\end{figure}

In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases. 
For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

@@

\paragraph{Parsing}
Our parsing workflow is multistep and primarily based on regular expressions.
In the case of instruction-tuned models, we first identify the \texttt{[ANSWER]}\texttt{[\textbackslash ANSWER]} environment we prompt the model to report the answer in.
In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
In the case of numbers, our regular expression was engineered to deal with various forms of scientific notation.
As initial tests indicated that models sometimes return integers in the form of words, e.g., \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., Claude 2, to parse the completion.

@@

\subparagraph{Instruction-tuned models} In addition, we used Claude 2, Claude3 (Opus),\autocite{anthropicClaudeModelFamily2024} GPT-4,\autocite{openai2024gpt4} GPT-3.5-turbo,\autocite{brown2020language} Gemini Pro,\autocite{gemini} Mixtral-8x7b\autocite{jiang2024mixtral} LLaMA2 (7B, 13B, 70B),\autocite{touvron2023llama} as well as the 7B chat model from Perplexity.AI.

@@

\paragraph{Question selection} \label{sec:subset-selection}
Since we anticipated that we would not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided to show a relevant subset of all questions to the human scorers.
For selecting the subset, we decided to address two questions:
\begin{itemize}
    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
    \item Are there areas in which the performance of humans is very different from the ones of the models?
\end{itemize}
To answer the first question, we selected up to 13 questions per source that \glspl{llm}did not answer correctly in an initial scoring round.
In addition, we picked 100 diverse ones using greedy MaxMin sampling on the embeddings on the questions computed using BART. We added five random questions about the number of \gls{nmr} signals and the point group of molecules.
% https://github.com/lamalab-org/chem-bench-app/issues/92 has the details
